---
title: "Reproducible Research: Peer Assessment 1"
output: html_document
---

Loading and preprocessing the data
```{r}
data <- read.csv("activity.csv")
data_waNA <- data[which(data$steps!= "NA"), ]
```
What is mean total number of steps taken per day?
```{r}
library(plyr)
daily_steps <- ddply(data_waNA, .(date), summarise, steps=sum(steps))
hist(daily_steps$steps, xlab="steps per day")
```
Mean number of steps per day is **`r mean(daily_steps$steps)`**  
Median number of steps per day is **`r median(daily_steps$steps)`**

What is the average daily activity pattern?
```{r}
average_date <- ddply(data_waNA, .(interval), summarise, steps=mean(steps))
plot(average_date$interval, average_date$steps, type="l", xlab="5-minute interval", 
ylab="Average steps",main="Average daily activity")

#Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
average_date[average_date$steps==max(average_date$steps),]
colnames(average_date)[2] <- "intervalAvg"
```
## Imputing missing values
```{r}
#  missing values
sum(is.na(data$steps))
# Imputing NA's with average on 5-min interval
merged <- arrange(join(data, average_date), interval)
# the new dataset 
merged$steps[is.na(merged$steps)] <- merged$intervalAvg[is.na(merged$steps)]
# plot the histogram
new_daily_steps <- ddply(merged, .(date), summarise, steps=sum(steps))
hist(new_daily_steps$steps, main="Number of Steps", 
     xlab="steps taken each day",,)

mean(new_daily_steps$steps)
median(new_daily_steps$steps)
daily_steps_1 <- sum(data_waNA$steps)
daily_steps_2 <- sum(merged$steps)
diff <- daily_steps_2 -daily_steps_1 []
```
Mean values didn't change

## Are there differences in activity patterns between weekdays and weekends?
```{r}
library(lattice)
weekdays <- weekdays(as.Date(merged$date))
data_weekdays <- transform(merged, day=weekdays)
data_weekdays$wk <- ifelse(data_weekdays$day %in% c("Saturday", "Sunday"),"weekend", "weekday")
average_week <- ddply(data_weekdays, .(interval, wk), summarise, steps=mean(steps))

xyplot(steps ~ interval | wk, data = average_week, layout = c(1, 2), type="l")
```
  
The graph shows that the patterns do not differ a lot between weekend and weekdays